Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2023-06-06 Manuscript_Colloids.tex
Add files via upload

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomencl}
\makenomenclature
\renewcommand{\nomlabel}[1]{\hfil #1\hfil}

@@

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection, further characterisation of trapped entities using light-scattering
  poses a two-fold challenge --- one experimental, concerning the optimal
  arrangement of light detectors to gather data, and the other theoretical, involving
  solving of the inverse light scattering problem in order to interpret this data. Experimentally, combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose here a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data, combined with a neural-network trained on data simulated to mimic light scattering
  detection signals in such scenarios, may be used for solving the
  inverse light scattering problem and characterising complex trapped
  entities.  To demonstrate the method we discuss its application to
  measuring the instantaneous orientations of a trapped asymmetric microsphere
  dimer. We argue that the method can be extended to
  determine any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

\begin{highlights}
\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.  
\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
\end{highlights}

@@

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}

Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico-newtons. Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.  Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass\cite{yifat2021facile}.  A recent work aimed at characterising trapped entities demonstrated how neural networks maybe trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}.  A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle. Thus, past studies on optical trapping have focused rather heavily on particle-tracking but not particle-characterisation.

While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics. A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion. While this provided dynamical information, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle. Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise isotropically scattering particles. This limits its application to biological matter as the internal structure of many cells makes them inherently isotropic scatterers \cite{Watson_2023}. In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation to provide maximal information from the signal.

To demonstrate the analysis, we study a simple, illustrative example of an isotropic scattering entity \emph{i.e.} an asymmetric dimer. As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate. As an example of the relatively sparse literature on measuring orientations of complex trapped objects, ref. employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted. Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer. We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals. Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
\end{subfigure}
\caption{\label{fig:setup}
  %
  Proposed experimental set up for scattering measurements from an object in an optical trap. The probe beam for scattering measurements is incident perpendicular to the trapping laser propagation direction. a) Side view. b) Top view. Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see text).
%
}
\end{figure}

\printnomenclature[0.55in]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}

\subsection{Brownian Simulation}
\label{sec:brownian}

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a 
highly focused Gaussian beam by calculating the optical forces imparted by 
the laser, and the Brownian forces imparted by the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
\begin{align}
	\begin{pmatrix}
		F_x \\ F_y \\ F_z \\ T_x \\ T_y \\ T_z 
	\end{pmatrix}
	= \begin{pmatrix}
	a_1 &   0  &     0     &   0  & d_1 &     0 \\
	 0  &  a_1 &     0     & -d_1 &  0  &     0 \\
	 0  &   0  & a_1 + a_2 &   0  &  0  &     0 \\
	 0  & -d_1 &     0     &  b_1 &  0  &     0 \\
	d_1 &   0  &     0     &   0  & b_1 &     0 \\
	 0  &   0  &     0     &   0  &  0  & b_1 + b_2     
	\end{pmatrix}
	\begin{pmatrix}
		v_{k0} - V_k \\ v_{k0} - V_k \\ v_{k0} - V_k \\
		\omega_x - \Omega_x \\ \omega_y - \Omega_y \\ \omega_z - \Omega_z
	\end{pmatrix}
\end{align}

where each matrix component is scaled by a factor of $k_B T/\pi \eta$ When both sphere's are identically sized - as in Fung~\textit{et. al}'s work - $d_1$ goes to zero; however for our work with asymmetric spheres these components are significant enough to be considered. We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation estimation from scattering measurements}
\label{sec:Bayes}

For our simulated dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere to the centre of the smaller sphere. A plane wave 'probe' laser, perpendicular to the trapping laser, is incident on the dimer, generating a scattering pattern  dependent on the dimer's orientation $I(\hat{\bf s}, \theta)$ which is computed using MSTM. To represent the experimental set up consisting of a set of optical fibres recording scattered light, we choose four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) and record the calculated intensity at each angle $\theta_k$, $I(\hat{s}, \theta_k)$.

Our goal is to determine the orientation of the trapped dimer based on the scattering data $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} \ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a unit sphere \cite{Rey2006} (Figure~\ref{fig:dimer}b) leading to a maximum nearest-neighbour spacing between two neighbouring reference orientations of 0.895 radians. Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
\label{eq:scale}
  y_k(\hat{\bf n}_\alpha)
  = 
  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}
where the denominator is simply the standard deviation across the set of values $I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex. Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle. 
Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks. We thus utilised the Python machine learning program \textit{scikit-learn} to build a neural network for identifying the dimer's orientation from its light scattering signal. The network was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then using the network to estimate the probability of a given signal coming from a dimer in a given reference orientation. The network's loss function was evaluated and used to improve the estimation, the network being trained until the improvement in the loss function was less than 0.0001. 
%
Importantly, the estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer,  utilising Bayesian inference to update the neural network's estimation:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimate's actual accuracy. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and  Discussion}
\label{sec:Discussion}
\subsection{Minimal number of detectors}
\label{sec:detectors}
The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200. For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres around a single point.

Initially we tried capturing scattered light only in the x-y plane as to simulate the detectors all being placed on a glass cover slip. The results of this estimate are shown below. As can be seen from Figure~ the neural network performs particularly poorly in this situation, seemingly jumping back and forth between different reference orientation at random. We found that when all three detectors are in the same plane the expected signal can appear identical despite the dimer being in completely different orientations. This is shown in Figure~ which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels shows that the dimer's scattering is indistinguishable in these two reference orientation. More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique. This is illustrative of the difficulty behind the inverse light scattering problem; as attributing the measured signal will not reveal any new information about the entity unless we can remove uncertainty around other aspects of the scatterer.

>>>>

\usepackage{csvsimple}
\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomencl}
\usepackage{tabularx}
\makenomenclature
\renewcommand{\nomlabel}[1]{\hfil #1\hfil}

@@

\begin{flushright}
	Daniel Maciver \\
	daniel.maciver.2016@uni.strath.ac.uk \\
	University of Strathclyde, \\
	75 Montrose Street, \\
	Glasgow, \\
	G1 1XL \\
	07793445441 
\end{flushright}

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection, further characterisation of trapped entities using light-scattering
  poses a two-fold challenge --- one experimental, concerning the optimal
  arrangement of light detectors to gather data and minimising signal noise, and the other theoretical, involving
  solving of the inverse light scattering problem in order to interpret this data. Experimentally, combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose here a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data, combined with a neural-network trained on data simulated to mimic light scattering
  detection signals in such scenarios, may be used for solving the
  inverse light scattering problem and characterising complex trapped
  entities.  To demonstrate the method we discuss its application to
  measuring the instantaneous orientations of a trapped asymmetric microsphere
  dimer. We argue that the method can be extended to
  determine any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

%\begin{highlights}
%\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.  
%\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
%\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
%\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
%\end{highlights}

@@

\end{frontmatter}
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:Intro}

Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico-newtons. Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.  Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass\cite{yifat2021facile}.  A recent work aimed at characterising trapped entities demonstrated how neural networks maybe trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}. A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle. Thus, past studies on optical trapping have focused on either on particle trapping to study trapping dynamics, or on the characterisation of particles with relatively simple complex trapping dynamics.

While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles with simple trapping dynamics, and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics. A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion - commonly now refereed to as Localised dynamic light scattering (LDLS). This was later expanded upon by \cite{PhysRevE.65.041921} by collecting back scattered light to characterise the Stokes friction coefficient as a function of trapping depth. While both papers provided dynamical information, structural information about the trapped bead was precluded as the scattered wave was only measured across a small angular range. Furthermore, the main drawback to a LDLS is that it struggles to characterise asymmetric particles such as nano rods or dimers; in which case the Brownian motion is no longer simply translational, but rotational about the centre of mass \cite{yin2013large}.

One way of mitigating this is to remove translational fluctuations from the analysis; by monitoring its the scattering, Cang and co-workers were able to spatially fix a gold nanorod at the centre of trapping laser by moving the sample plane accordingly, to an accuracy of 200 nm \cite{Cang2006}. However, this has limited application's to anisotropic scatters such as biological matter where the internal structure of many cells makes them inherently anisotropic \cite{Watson_2023}. Thus even if the Cartesian coordinates are relatively fixed characterisation of the scattering pattern will have to separate the contributions of internal structure and orientation. In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique, to enable interpretation of the resulting multi-angle data from an anisotropic, asymmetric scatterer, as well as optimisation to provide maximal information from the signal.

To demonstrate the analysis, we study a simple, illustrative example of an anisotropic scattering entity \emph{i.e.} an asymmetric dimer. As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate. As an example of the relatively sparse literature on measuring orientations of complex trapped objects, \cite{raudsepp2022estimating} employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted. Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer. We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals. Furthermore we demonstrate how the model's performance when dealing with signal noise, a common problem when analysing scattering behaviour.

\begin{figure} [b]
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
\end{subfigure}
\caption{\label{fig:setup}
  %
  Proposed experimental set up for scattering measurements from an object in an optical trap. The probe beam for scattering measurements is incident perpendicular to the trapping laser propagation direction. a) Side view. b) Top view. Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see Sec~\ref{sec:detectors}).
%
}
\end{figure}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}
\subsection{Orientation estimation from scattering measurements}
\label{sec:Bayes}

Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere to the centre of the smaller sphere. A plane wave 'probe' laser, perpendicular to the trapping laser, is incident on the dimer, generating a scattering pattern  dependent on the dimer's orientation $I(\hat{\bf s}, \theta)$ which can be computed using  software such as MSTM \cite{Mishchenko1996MSTM}. To represent the experimental set up consisting of a set of optical fibres recording scattered light, we choose four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) and record the calculated intensity at each angle $\theta_k$, $I(\hat{s}, \theta_k)$.

@@

Our goal is to determine the orientation of the trapped dimer based on the measured intensity $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} \ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a unit sphere \cite{Rey2006} (Figure~\ref{fig:dimer}b) leading to a maximum nearest-neighbour spacing between two neighbouring reference orientations of 0.895 radians. Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
\label{eq:scale}
  y_k(\hat{\bf n}_\alpha)
  = 
  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}
where the denominator is simply the standard deviation across the set of values $I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem. While it is trivial to compute the light scattering pattern for any given particle with any particular characteristic (i.e. size, shape, or orientation), inferring the light scattering from a unknown particle to determine said characteristic is incredibly difficult due to complex mapping between scattering and said characteristic. Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle. Shown below is two clusters of orientation vectors and there respective measured scattering signals - the points have been coloured based on their proximity to the centre of their respective cluster. While the orientation space appears tightly packed and ordered the signal space quickly spreads out in an asymmetric fashion. Furthermore as seen in Fig~\ref{fig:mixing}b the signal mapping can intersect itself which only further increases the complexity. While in some instances the mapping between one reference orientation and another is discrete, in other instances the mapping becomes far more complex to discern.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks. We thus utilised the Python machine learning program \textit{scikit-learn} to build a neural network for identifying the dimer's orientation from its light scattering signal. The network was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then using the network to estimate the probability of a given signal coming from a dimer in a given reference orientation. The network's loss function was evaluated and used to improve the estimation, the network being trained until the improvement in the loss function was less than 0.0001. 
%
Importantly, the estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer, utilising Bayesian inference to update the neural network's estimation:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brownian Simulation}
\label{sec:brownian}

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-3} \ s$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and  Discussion}
\label{sec:Discussion}
\subsection{Minimal number of detectors}
\label{sec:detectors}
The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200. For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.
